Robust calibration of numerical models based on relative regret

### 2019-11-20 article-modif-A&E.tex
autoformatting

In this paper, we aim at finding $\kest$ a {\it robust} estimator of $\mathbf{k}$.
The definition of robustness differs depending on the context in which it is used. Indeed, one definition of the robustness of an estimate is a measure of the sensibility of said estimate to outliers~\citep{huber_robust_2011}. This lead to the introduction of robust norms in data assimilation \citep{rao_robust_2015}. %The point of this norm is to assume Gaussian distribution for small errors (the usual $L_2$ norm), while the large errors are assumed to be Laplace distributed (so the $L_1$ norm).
In a Bayesian framework, robustness may refer to the sensitivity to a wrong specification of the priors~\citep{berger_overview_1994}.
%
Throughout this paper, 
%we define robustness as the ability of $\kest$ to perform reasonably well under different operating conditions (i.e.\ different $\mathbf{u}$).
robust has to be understood as satisfactory for a broad range of $\mathbf{u}$, and/or as insensitive as possible to uncertainties encompassed in $\mathbf{U}$.

@@

The random variable $\mathbf{U}$ is assumed to be uniformly distributed over $\Uspace$.
%The definition and the estimation of $\kmean$ and $\kvar$ is pretty standard.  
The estimations are %$\mathbf{k}\mapsto \mu(\mathbf{k})$ and $\mathbf{k}\mapsto\sigma(\mathbf{k})$ 
based on a $1000 \times 1000$ regular grid over $\Kspace\times\Uspace$. Both cost functions are shown in  Figure~\ref{fig:branin_moments}, top row.

@@

Given that $J>0$, $\Gamma_{\alpha}(\mathbf{k})$ is increasing with respect to $\alpha$ for any $\mathbf{k}\in\Kspace$. We can then focus on the smallest value of $\alpha$ such that $\Gamma_\alpha$ reaches a certain level of confidence $p\in[0,1]$. This leads to the definition of $\checkap$
\begin{align}
  \checkap &= \inf\left\{ \alpha\geq 1 \mid \exists \checkkp \in \Kspace,\, \Gamma_{\alpha}(\checkkp) \geq p \right\} \nonumber \\
   &= \inf \left\{ \alpha \geq 1 \mid \max_{\mathbf{k}\in\Kspace} \Gamma_{\alpha}(\mathbf{k}) \geq p \right\}   \label{eq:def_alpha_check}
\end{align}
that is the smallest $\alpha$, such that there exists a particular $\checkkp \in \Kspace$ for which $J(\checkkp,\mathbf{U}) \leq \checkap J^*(\mathbf{U})$ with probability $p$. This formulation can be linked to the quantiles and Value-at-Risk of the random variable $\max_{\mathbf{k}} \{ \frac{J(\mathbf{k},\mathbf{U})}{J^*(\mathbf{U})}\}$, that is a measure of risk usually applied in the financial sector (see~\cite{rockafellar_deviation_2002}). For different levels $p$, and thus different $\alpha_p$, Figure~\ref{fig:illu_alpha_p} shows examples of $\Gamma_{\alpha_p}$. By definition, the associated $\mathbf{k}_p$ is then the first value for which $\Gamma_{\alpha_p}$ reaches $p$. We can see that changing the level $p$ shifts the maximiser of $\Gamma_{\alpha_p}$, and that for small $\alpha$, $\Gamma_{\alpha}(\mathbf{k}_1)$ is also very small. This indicates that $\mathbf{k}_1$ is located quite far from the conditional minimisers, and thus represents a compromise.

@@

%We are now going to compute and compare those estimators on the $J_{\mathrm{BH}}$and $J_{\mathrm{BHswap}}$ function.

@@

\begin{align}
  \label{eq:SWE_equations}
  \left\{
    \begin{array}{rl}
      \partial_t h + \partial_x q &= 0 \\
      \partial_t q + \partial_x \left(\frac{q^2}{h} + \frac{g}{2}h^2\right) &= - gh\partial_x z - \mathbf{k}q |q|h^{-7/3} \\
      h(0, t) &= 20.0 + 3\cdot \sin \left(\frac{2\pi t}{2}\right) + 1.5\cdot \mathbf{u} \cdot \sin\left(\frac{2 \pi t}{\omega_0}\right)\\
      \partial_x q(0, t) &= 0
      \end{array}
  \right.
\end{align}
These equations are integrated using a finite-volume scheme on a discretized domain $[0, L]$, up to a time $T$. The output of the computer code is the sea surface height $h$, on the center of all the volumes and at all the time-steps, that will be denoted $\mathcal{M}(\mathbf{k},\mathbf{u};\, \omega_0=1.0)$.
In this setting, the variable input is uniformly distributed on $\mathcal{U}$.
To generate the observations, we choose $\mathbf{u}^{\mathrm{true}}=2/3$, and define $\mathcal{M}^o$ based on the computer model $\mathcal{M}$, such that $\mathcal{M}^o(\mathbf{u}^{\mathrm{true}}) = \mathcal{M}(\mathbf{k}^{\mathrm{true}}, \mathbf{u}^{\mathrm{true}}; \, \omega_0 = 0.999)$. $\omega_0$  
represents here the uncontrollable error between the observations and the numerical model and will now be omitted systematically in the notation. The true value of the bottom friction $\mathbf{k}^{\mathrm{true}} = (k_1^{\mathrm{true}}, k_2^{\mathrm{true}},\dots, k_{N_\mathrm{vol}}^{\mathrm{true}})$ is not constant over the whole domain, and is defined as
\begin{equation*}
{k}^{\mathrm{true}}_i = 0.2\cdot \left(1 + \sin\left(\frac{2 \pi x_i}{L}\right)\right)
\end{equation*}
where $x_i$ is the center of the $i$-th volume. The two sources of representativity errors are $\omega_0$, and the fact that we are looking to estimate $\mathbf{k}$ in a 1-dimensional space $\Kspace$. 
\subsection{Estimation}
Given that the numerical model is expensive to evaluate, it has first been evaluated on a relatively small regular grid. A metamodel based on Gaussian process is constructed and the PEI criterion \citep{ginsbourger_bayesian_2014,bossek_learning_2015} allow us to add point to evaluate and to better capture the locus of the conditional minimisers $\{(\mathbf{k}^*(\mathbf{u}), \mathbf{u}) \mid \mathbf{u} \in \Uspace \}$. A bigger regular grid is evaluated by the metamodel once the design space has been sufficiently explored. The different steps of the estimation are illustrated Figure~\ref{fig:estimation_swe}. We can see that the estimation for this problem seems less problematic than the analytical examples shown above, as all derived quantites are unimodal.
The different estimates however take a wide range of values, as seen on Table~\ref{tab:recap_estimates_metaSWE}. As a basis for comparison, the global minimiser of $J$ over $\Kspace \times \Uspace$ has been computed, and $\kest_{\mathrm{global}}$ is then obtained by discarding the $\mathbf{u}$ value. An interesting comparison arises when comparing with $\kest_{\mathrm{global}}$, where $\kest_1$ outperforms the global minimiser for $\mathbf{u} < 0.625$.

>>>>

In this paper, we aim at finding $\kest$ a {\it robust\/} estimator of $\mathbf{k}$.
The definition of robustness differs depending on the context in which it is used. Indeed, one definition of the robustness of an estimate is a measure of the sensibility of said estimate to outliers~\citep{huber_robust_2011}. This lead to the introduction of robust norms in data assimilation \citep{rao_robust_2015}. %The point of this norm is to assume Gaussian distribution for small errors (the usual $L_2$ norm), while the large errors are assumed to be Laplace distributed (so the $L_1$ norm).
In a Bayesian framework, robustness may refer to the sensitivity to a wrong specification of the priors~\citep{berger_overview_1994}.
%
Throughout this paper, 
%we define robustness as the ability of $\kest$ to perform reasonably well under different operating conditions (i.e.\ different $\mathbf{u}$).
robust has to be understood as satisfactory for a broad range of $\mathbf{u}$, and/or as insensitive as possible to uncertainties encompassed in $\mathbf{U}$.

@@

The random variable $\mathbf{U}$ is assumed to be uniformly distributed over $\Uspace$.
%The definition and the estimation of $\kmean$ and $\kvar$ is pretty standard.  
The estimations are %$\mathbf{k}\mapsto \mu(\mathbf{k})$ and $\mathbf{k}\mapsto\sigma(\mathbf{k})$ 
based on a $1000 \times 1000$ regular grid over $\Kspace\times\Uspace$. Both cost functions are shown on~Figure~\ref{fig:branin_moments}, top row.

@@

Given that $J>0$, $\Gamma_{\alpha}(\mathbf{k})$ is increasing with respect to $\alpha$ for any $\mathbf{k}\in\Kspace$. We can then focus on the smallest value of $\alpha$ such that $\Gamma_\alpha$ reaches a certain level of confidence $p\in[0,1]$. This leads to the definition of $\checkap$
\begin{align}
  \checkap &= \inf\left\{ \alpha\geq 1 \mid \exists \checkkp \in \Kspace,\, \Gamma_{\alpha}(\checkkp) \geq p \right\} \nonumber \\
   &= \inf \left\{ \alpha \geq 1 \mid \max_{\mathbf{k}\in\Kspace} \Gamma_{\alpha}(\mathbf{k}) \geq p \right\}   \label{eq:def_alpha_check}
\end{align}
that is the smallest $\alpha$, such that there exists a particular $\checkkp \in \Kspace$ for which $J(\checkkp,\mathbf{U}) \leq \checkap J^*(\mathbf{U})$ with probability $p$. This formulation can be linked to the quantiles and Value-at-Risk of the random variable $\max_{\mathbf{k}} \{ \frac{J(\mathbf{k},\mathbf{U})}{J^*(\mathbf{U})}\}$, that is a measure of risk usually applied in the financial sector (see~\cite{rockafellar_deviation_2002}). For different levels $p$, and thus different $\alpha_p$, Figure~\ref{fig:illu_alpha_p} shows examples of $\Gamma_{\alpha_p}$. By definition, the associated $\mathbf{k}_p$ is then the first value for which $\Gamma_{\alpha_p}$ reaches $p$. We can see that changing the level $p$ shifts the maximiser of $\Gamma_{\alpha_p}$, and that for small $\alpha$, $\Gamma_{\alpha}(\mathbf{k}_1)$ is also very small. This indicates that $\mathbf{k}_1$ is located quite far from the conditional minimisers, and arise as a compromise when the relaxation is large enough.

@@

% We are now going to compute and compare those estimators on the $J_{\mathrm{BH}}$and $J_{\mathrm{BHswap}}$ function.

@@

\begin{align}
  \label{eq:SWE_equations}
  \left\{
    \begin{array}{rl}
      \partial_t h + \partial_x q &= 0 \\
      \partial_t q + \partial_x \left(\frac{q^2}{h} + \frac{g}{2}h^2\right) &= - gh\partial_x z - \mathbf{k}q |q|h^{-7/3} \\
      h(0, t) &= 20.0 + 3\cdot \sin \left(\frac{2\pi t}{2}\right) + 1.5\cdot \mathbf{u} \cdot \sin\left(\frac{2 \pi t}{\omega_0}\right)\\
      \partial_x q(0, t) &= 0
      \end{array}
  \right.
\end{align}
These equations are integrated using a finite-volume scheme on a discretized domain $[0, L]$, up to a time $T$. The output of the computer code is the sea surface height $h$, on the center of all the volumes and at all the time-steps, that will be denoted $\mathcal{M}(\mathbf{k},\mathbf{u};\, \omega_0=1.0)$.
In this setting, the variable input is uniformly distributed on $\mathcal{U}$.
To generate the observations, we choose $\mathbf{u}^{\mathrm{true}}=2/3$, and define $\mathcal{M}^o$ based on the computer model $\mathcal{M}$, such that $\mathcal{M}^o(\mathbf{u}^{\mathrm{true}}) = \mathcal{M}(\mathbf{k}^{\mathrm{true}}, \mathbf{u}^{\mathrm{true}}; \, \omega_0 = 0.999)$. $\omega_0$  
represents here the uncontrollable error between the observations and the numerical model and will now be omitted systematically in the notation. The true value of the bottom friction $\mathbf{k}^{\mathrm{true}} = (k_1^{\mathrm{true}}, k_2^{\mathrm{true}},\dots, k_{N_\mathrm{vol}}^{\mathrm{true}})$ is not constant over the whole domain, and is defined as
\begin{equation*}
{k}^{\mathrm{true}}_i = 0.2\cdot \left(1 + \sin\left(\frac{2 \pi x_i}{L}\right)\right)
\end{equation*}
where $x_i$ is the center of the $i$-th volume. The two sources of representativity errors are $\omega_0$, and the fact that we are looking to estimate $\mathbf{k}$ in a 1-dimensional space $\Kspace$. 
\subsection{Estimation}
Given that the numerical model is expensive to evaluate, it has first been evaluated on a relatively small regular grid. A metamodel based on Gaussian process is constructed and the PEI criterion~\citep{ginsbourger_bayesian_2014,bossek_learning_2015} allow us to add point to evaluate and to better capture the locus of the conditional minimisers $\{(\mathbf{k}^*(\mathbf{u}), \mathbf{u}) \mid \mathbf{u} \in \Uspace \}$. A bigger regular grid is evaluated by the metamodel once the design space has been sufficiently explored. The different steps of the estimation are illustrated Figure~\ref{fig:estimation_swe}. We can see that the estimation for this problem seems less problematic than the analytical examples shown above, as all derived quantites are unimodal.
The different estimates however take a wide range of values, as seen on Table~\ref{tab:recap_estimates_metaSWE}. As a basis for comparison, the global minimiser of $J$ over $\Kspace \times \Uspace$ has been computed, and $\kest_{\mathrm{global}}$ is then obtained by discarding the $\mathbf{u}$ value.
